---
title: "Data Wrangling Exersize 1"
author: "r.b.metcalf"
date: "9 Aug 2016"
output: html_document
---

```{r,echo=FALSE,message=FALSE}
library("tidyr")
library("dplyr")
```



## Data Wrangling Exercise 1

The data file was converted into csv format with OpenOffice.

Now we load it into R and take a look at the structure.
```{r}
table <- tbl_df(read.csv("refine_original.csv"))
glimpse(table)
```

Find the unique company names and clean them up.

```{r}
table %>% distinct(company)

table$company <- tolower(table$company)
table$company <- gsub("0","o",table$company)
table$company <- gsub("phillips","philips",table$company)
table$company <- gsub("phllips","philips",table$company)
table$company <- gsub("phillps","philips",table$company)
table$company <- gsub("fillips","philips",table$company)
table$company <- gsub("phlips","philips",table$company)
table$company <- gsub("ak zo","akzo",table$company)
table$company <- gsub("unilver","unilever",table$company)

table %>% distinct(company)
```

Now separate product code and number into different columns.
```{r}
table <- table %>% mutate(Product.code = Product.code...number) %>% mutate(Product.numbers = Product.code...number)

table$Product.code <- gsub("\\-.*","",table$Product.code)
table$Product.numbers <- as.numeric(gsub(".*\\-","",table$Product.numbers))
```

*Now convert the product codes to real names.*

Define a function that converts codes to names.
```{r}
product_name <- function(x){
  if(x == "p") return("Smartphone")
  if(x == "v") return("TV")
  if(x == "x") return("Laptop")
  if(x == "q") return("Tablet")
  return("NA")  
}
```
Now make new column with names Product.name
```{r}
table$Product.name <- sapply(table$Product.code,product_name)
glimpse(table)
```
Concatenate address parts
```{r}
table$full_address <- paste(table$city,",",table$address,",",table$country)
table$full_address <- gsub(" ,",",",table$full_address)
```
Add columns for binary company classification
```{r}
table$company_philips <- table$company == "philips"
table$company_akzo <- table$company == "akzo"
table$company_van_houten <- table$company == "van houten"
table$company_unilever <- table$company == "unilever"

## check to see that all of them have a classification
table %>% select(company_philips:company_unilever) %>% rowSums()

```

Add columns for binary product classification
```{r}
table$product_smartphone <- table$Product.name == "Smartphone"
table$product_tv <- table$Product.name == "TV"
table$product_laptop <- table$Product.name == "Laptop"
table$product_tablet <- table$Product.name == "Tablet"

## check to see that all of them have a classification

table %>% select(product_smartphone:product_tablet) %>% rowSums()
```

```{r}
table %>% write.table("refine_clean.csv")
```

